---
title: "Lab 3a - Exercises"
author: "Environmental Data Analytics | John Fay and Luana Lima "
date: "Fall 2024"
output: pdf_document
---

## Set up the session
```{r Setup}
#Import packages
library(tidyverse); library(lubridate); library(here)

#Check workspace
here()

#Import data
pm25 <- read.csv(
  here('./Data/Raw/EPAair_PM25_NC2019_raw.csv'),
  stringsAsFactors = TRUE,
  colClasses = c("Site.ID"='factor')
)

#Fix the date
pm25$Date <- mdy(pm25$Date)

```

## Explore properties of the dataframe
-   How many different sites are included in the data set?
-   What is the mean daily PM2.5 concentration?
-   In what units of are PM 2.5 concentrations reported?

```{r Explore the data}
# Compute the number of different sites
summary(pm25$Site.ID)  #-or-
unique(pm25$Site.ID)  #-or-
levels(pm25$Site.ID)  #-or-
length(summary(pm25$Site.ID))  #-or-
length(unique(pm25$Site.ID))  #-or-
length(levels(pm25$Site.ID))

# Compute the mean and median PM 2.5 concentration
mean(pm25$Daily.Mean.PM2.5.Concentration)  #-or-
median(pm25$Daily.Mean.PM2.5.Concentration)  #-or-
summary(pm25$Daily.Mean.PM2.5.Concentration)

# What are the units
summary(pm25$UNITS)

# What is the date range (min and max dates)
summary(pm25$Date)  #-or-
min(pm25$Date)  #-or-
max(pm25$Date)
```

## Plot distribution of values

-   Number of samples using each different source 
-   Distribution of daily mean concentration values
-   Distribution of daily mean concentration, by source

```{r}
#Number of samples using each different source 
ggplot(pm25, aes(x=Source)) +   geom_bar()

#Distribution of daily mean concentration values
ggplot(pm25, aes(x=Daily.Mean.PM2.5.Concentration)) +   geom_histogram()

#Distribution of daily mean concentration, by source
##Histogram
ggplot(pm25) + 
  geom_histogram(aes(x=Daily.Mean.PM2.5.Concentration, fill=Source))

##Frequency Polygon
ggplot(pm25) + 
  geom_freqpoly(aes(x=Daily.Mean.PM2.5.Concentration, color=Source))

##Box plot
ggplot(pm25, aes(y=Daily.Mean.PM2.5.Concentration, x=Source))  +  
  geom_boxplot()

##Violin plot
ggplot(pm25, aes(y=Daily.Mean.PM2.5.Concentration, x=Source))  +  
  geom_violin()

```

